```{r} 
# Load necessary libraries 
library("tximport") 
library("readr")
library("DESeq2")
library("tximportData")
library("ggplot2")
```

```{R}
# List all of the quant.sf files
cwd <- getwd() 
quant_fp <- paste(cwd, 'quants', sep = '/') 
quant_files<- list.files(path = quant_fp, pattern = "quant.sf$", recursive = TRUE)
```

```{r}
# Extract sample names
quant_dirs <- list.files(quant_fp, 
                        pattern = "_quant$", 
                        full.names = TRUE)
sample_names <- gsub("_quant$", "", basename(quant_dirs)) 
# Correct for the FASTQ generation file included in the directory sample_names \<-
sample_names <- sample_names[1:48] # Ensure this range fits your data
```

```{r}
quant_dirs
```

```{r}
# Create the txi import
txi <- tximport(files = paste(quant_fp, quant_files, sep = '/'), type = "salmon", txOut = TRUE)
```

```{r}
colnames(txi$abundance) <- sample_names
colnames(txi$counts) <- sample_names
colnames(txi$length) <- sample_names
```

```{r}
runDESeq2 <- function(sample_names, txi, cond1, cond2){
sample_info <- data.frame(
    sample_names = sample_names,
    condition = factor(ifelse(grepl(cond1, sample_names), cond1, 
                              ifelse(grepl(cond2, sample_names), cond2, NA)))
)
sample_info <- sample_info[!is.na(sample_info$condition), ]

samples_to_keep <- sample_info$sample_names
print(samples_to_keep)

# Subset txi
txi_filtered <- list(
  counts = txi$counts[, samples_to_keep],
  abundance = txi$abundance[, samples_to_keep],
  length = txi$length[, samples_to_keep],
  countsFromAbundance = "scaledTPM"
)
dds <- DESeqDataSetFromTximport(txi = txi_filtered, colData = sample_info, design = ~condition)
return(dds)
}
dds <- runDESeq2(sample_names, txi, "NC_1", "NC_ind_1")
```

```{r}
dds <- DESeq(dds)
```
```{r}
res <- results(dds)
```

```{r}
plotMA(res)
```
```{r}
dds2 = runDESeq2(sample_names, txi, "WT_1","NC_1" )
```

```{r}
dds2 <- DESeq(dds2)
res2 <- results(dds2)
```
```{r}
plotMA(res2)
```
```{r}
condition_1 = matrix(c("WT_1", "WT_ind_1", 
                       "WT_1", "NC_1", 
                       "WT_1", "NC_ind_1",
                       "WT_1", "s4_1",
                       "WT_1", "s4_ind_1"),
                       nrow=5,
                       ncol=2,
                       byrow=TRUE)
print(condition_1)
```
```{r}
dds1_list = list()
for(row in 1:nrow(condition_1)){
  cond1 = condition_1[row,1]
  cond2 = condition_1[row,2]
  dds = runDESeq2(sample_names, txi, cond1, cond2)
  dds1_list = append(dds_list, list(dds))
}
print(dds1_list)
```
```{r}
results1_list = list()
for(i in 1:length(dds1_list)){
  dds = dds1_list[[i]]
  dds = DESeq(dds)
  result = results(dds)
  results1_list = append(results1_list, list(result))
  
  res <- as.data.frame(results(dds))
  
  timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
  filename <- paste0("deseq2Timepoint1_", timestamp, "_", condition_1[i,1], "_", condition_1[i,2], ".csv")
  write.csv(res, file = filename, row.names=TRUE)
  
  res$significance <- ifelse(res$padj < 0.05 & abs(res$log2FoldChange) > 1, "Significant", "Not Significant")

# Create the volcano plot
# Create the volcano plot without the title, with legend inside, and larger global font size
  print(ggplot(res, aes(x = log2FoldChange, y = -log10(padj), color = significance)) +
       geom_point(alpha = 0.6, size = 1.5) +
        scale_color_manual(values = c("gray", "red")) +
        theme_minimal(base_size = 14) +  # Set a base global font size (adjust as needed)
        geom_vline(xintercept = c(-1, 1), col = "blue", linetype = "dashed") +
        geom_hline(yintercept = -log10(0.05), col = "blue", linetype = "dashed") +
        labs(title = NULL,  # Remove the title
             x = "Log2 Fold Change", y = "-Log10 Adjusted P-Value") +  # Adjust text size for labels  # Adjust the y-axis range # Specify breaks for y-axis ticks
        theme(
          legend.position = c(0.89, 0.5),  # Move the legend inside the plot on the right side
          legend.title = element_text(size = 14),  # Increase the font size of the legend title
          legend.text = element_text(size = 12),  # Increase the font size of the legend text
          plot.title = element_blank(),  # Ensure no title
          text = element_text(size = 14),  # Global font size adjustment for the entire plot
          legend.background = element_rect(color = "black", fill = NULL)  # Add a black border around the legend
        )
  )
}
```
```{r}
condition_3 = matrix(c("WT_3", "WT_ind_3", 
                       "WT_3", "NC_3", 
                       "WT_3", "NC_ind_3",
                       "WT_3", "s4_3",
                       "WT_3", "s4_ind_3"),
                       nrow=5,
                       ncol=2,
                       byrow=TRUE)
print(condition_3)
```
```{r}
dds3_list = list()
for(row in 1:nrow(condition_3)){
  cond1 = condition_3[row,1]
  cond2 = condition_3[row,2]
  dds = runDESeq2(sample_names, txi, cond1, cond2)
  dds3_list = append(dds3_list, list(dds))
}
print(dds3_list)
```

```{r}
results3_list = list()
for(i in 1:length(dds3_list)){
  dds = dds3_list[[i]]
  dds = DESeq(dds)
  result = results(dds)
  results2_list = append(results3_list, list(result))
  
  res <- as.data.frame(results(dds))
  ## Save timepoint one data to a csv.
  timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
  filename <- paste0("deseq2Timepoint3_", timestamp, "_", condition_3[i,1], "_", condition_3[i,2], ".csv")
  write.csv(res, file = filename, row.names=TRUE)

  res$significance <- ifelse(res$padj < 0.05 & abs(res$log2FoldChange) > 1, "Significant", "Not Significant")

# Create the volcano plot
# Create the volcano plot without the title, with legend inside, and larger global font size
  print(ggplot(res, aes(x = log2FoldChange, y = -log10(padj), color = significance)) +
       geom_point(alpha = 0.6, size = 1.5) +
        scale_color_manual(values = c("gray", "red")) +
        theme_minimal(base_size = 14) +  # Set a base global font size (adjust as needed)
        geom_vline(xintercept = c(-1, 1), col = "blue", linetype = "dashed") +
        geom_hline(yintercept = -log10(0.05), col = "blue", linetype = "dashed") +
        labs(title = NULL,  # Remove the title
             x = "Log2 Fold Change", y = "-Log10 Adjusted P-Value") +  # Adjust text size for labels  # Adjust the y-axis range # Specify breaks for y-axis ticks
        theme(
          legend.position = c(0.89, 0.5),  # Move the legend inside the plot on the right side
          legend.title = element_text(size = 14),  # Increase the font size of the legend title
          legend.text = element_text(size = 12),  # Increase the font size of the legend text
          plot.title = element_blank(),  # Ensure no title
          text = element_text(size = 14),  # Global font size adjustment for the entire plot
          legend.background = element_rect(color = "black", fill = NULL)  # Add a black border around the legend
        )
  )
}
```
```{r}
library(tximeta)
sample_names
sample_info
files <- file.path(quant_fp, quant_files)
quant_fp
files
```
```{r}
coldata <- sample_info
coldata
coldata$names <- paste(coldata$names, "quant", sep="_")
coldata$files <- files
files
file.exists(coldata$files)
```

```{r}
splitstrs <- strsplit(sample_names, "\\_|\\-")
string = splitstrs[[1]]
length(splitstrs)

```

```{r}
strain<-c()
treatment<-c()
day<-c()
replicate=c()
for (i in 1:length(sample_names)){
  strippedConds = strsplit(sample_names, "\\_|\\-")
  strain <- append(strain, strippedConds[[i]][2])
  if (length(strippedConds[[i]]) == 4){
    treatment <- append(treatment, "untreated")
    day <- append(day, strippedConds[[i]][3])
    replicate <- append(replicate, strippedConds[[i]][4])
  }
  else if (length(strippedConds[[i]]) == 5){
    treatment = append(treatment, "treated")
    day <- append(day, strippedConds[[i]][4])
    replicate <- append(replicate, strippedConds[[i]][5])
  }
  
}
```
```{r}
strain
treatment 
day
replicate
```
```{r}
si_complete <-data.frame(samples=sample_names,
                         strain=strain,
                         treatment=treatment,
                         day=day,
                         replicate=replicate)
```
```{r}
si_complete
```
```{r}
coldata
```
```{r}
coldata <- si_complete
coldata
coldata$names <- coldata$samples
coldata$files <- files
coldata
```
```{r}
se <- tximeta(coldata) #SummarizedExperiment Object
```
```{r}
head(assay(se), 3)
```
```{r}
colSums(assay(se))
```
```{r}
se$treatment
```
```{r}
library(magrittr)
if (!is.factor(se$treatment)){
  se$treatment <- factor(se$treatment)
}
se$treatment %<>% relevel("untreated")
se$day 

```
```{r}
se$treatment

```
```{r}
head(se$treatment)
```

```{r}
round(colSums(assay(se))/1e6, 3)

```
```{r}
dds <- DESeqDataSet(se, design = ~ replicate + strain + treatment)

```
# Run Prefiltering because DESeqDataSet contains several rows that contain zeros or small amounts. Prefiltering will remove these and allow for increased speed when running functions.
```{r}
nrow(dds)
smallestGroupSize <- 4
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]
nrow(dds)


```

# Lets look at the simulated data of what the counts should look like if they follow a poisson distribution.
```{r}
lambda <- 10^seq(from = -1, to = 2, length = 1000)
cts <- matrix(rpois(1000*100, lambda), ncol = 100)
library("vsn")
meanSdPlot(cts, ranks = FALSE)

log.cts.one <- log2(cts + 1)
meanSdPlot(log.cts.one, ranks = FALSE)
```
# Now we will compare the variance stabilizing transformation (VST), regularized-logarithm transformation (rlog), and log2 transformation.
```{r}
vsd <- varianceStabilizingTransformation(dds, blind = FALSE)
head(assay(vsd), 3)
```
```{r}
rld <- rlog(dds, blind = FALSE)
head(assay(rld), 3)
```
```{r}
library("dplyr")
library("ggplot2")

dds <- estimateSizeFactors(dds)

df <- bind_rows(
  as_data_frame(log2(counts(dds, normalized=TRUE)[, 1:2]+1)) %>%
         mutate(transformation = "log2(x + 1)"),
  as_data_frame(assay(vsd)[, 1:2]) %>% mutate(transformation = "vst"),
  as_data_frame(assay(rld)[, 1:2]) %>% mutate(transformation = "rlog"))
  
colnames(df)[1:2] <- c("x", "y")  

lvls <- c("log2(x + 1)", "vst", "rlog")
df$transformation <- factor(df$transformation, levels=lvls)

ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed() + facet_grid( . ~ transformation) 
```
# Now we can look at the euclidean distance between the sample to check for similarity between them.
```{r}
sampleDists <- dist(t(assay(vsd)))
sampleDists
```
# And Create a heatmap to visualize.
```{r}
library("pheatmap")
library("RColorBrewer")
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( vsd$treatment, vsd$strain, sep = " - " )
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
```
# Do the same using a poisson distance.
```{r}
library("PoiClaClu")
poisd <- PoissonDistance(t(counts(dds)))
samplePoisDistMatrix <- as.matrix( poisd$dd )
rownames(samplePoisDistMatrix) <- paste( dds$treatment, dds$strain, sep=" - " )
colnames(samplePoisDistMatrix) <- NULL
pheatmap(samplePoisDistMatrix,
         clustering_distance_rows = poisd$dd,
         clustering_distance_cols = poisd$dd,
         col = colors)

```
# Finally with a PCA.
```{r}
plotPCA(vsd, intgroup = c("treatment", "strain"))
```
```{r}
pcaData <- plotPCA(vsd, intgroup = c( "treatment", "strain"), returnData = TRUE)
pcaData
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = treatment, shape = strain)) +
  geom_point(size =3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() +
  ggtitle("PCA with VST data")
```
# Finally, finally with a generalized PCA (GLM-PCA)
```{r}
library("glmpca")
gpca <- glmpca(counts(dds), L=2)
gpca.dat <- gpca$factors
gpca.dat$dex <- dds$treatment
gpca.dat$cell <- dds$strain

ggplot(gpca.dat, aes(x = dim1, y = dim2, color = dex, shape = cell)) +
  geom_point(size =3) + coord_fixed() + ggtitle("glmpca - Generalized PCA")
```
# multidimensional scaling plot.
```{r}
mds <- as.data.frame(colData(vsd))  %>%
         cbind(cmdscale(sampleDistMatrix))
ggplot(mds, aes(x = `1`, y = `2`, color = treatment, shape = strain)) +
  geom_point(size = 3) + coord_fixed() + ggtitle("MDS with VST data")
```
```{r}
dds <- DESeq(dds)
res <- results(dds)
res
```
```{r}
topGene <- rownames(res)[which.min(res$padj)]
plotCounts(dds, gene = topGene, intgroup=c("treatment"))
```
```{r}
library("ggbeeswarm")
geneCounts <- plotCounts(dds, gene = topGene, intgroup = c("treatment","strain"),
                         returnData = TRUE)
ggplot(geneCounts, aes(x = treatment, y = count, color = strain)) +
  scale_y_log10() +  geom_beeswarm(cex = 3)
```
```{r}
ggplot(geneCounts, aes(x = treatment, y = count, color = strain, group = strain)) +
  scale_y_log10() + geom_point(size = 3) + geom_line()
```
```{r}
library("apeglm")
resultsNames(dds)

res <- lfcShrink(dds, coef="treatment_treated_vs_untreated", type="apeglm")
plotMA(res, ylim = c(-5, 5))
res.noshr <- results(dds, name="treatment_treated_vs_untreated")
plotMA(res.noshr, ylim = c(-5, 5))
plotMA(res, ylim = c(-5,5))
topGene <- rownames(res)[which.min(res$padj)]
with(res[topGene, ], {
  points(baseMean, log2FoldChange, col="dodgerblue", cex=2, lwd=2)
  text(baseMean, log2FoldChange, topGene, pos=2, col="dodgerblue")
})
hist(res$pvalue[res$baseMean > 1], breaks = 0:20/20,
     col = "grey50", border = "white")
```
```{r}
library("genefilter")
topVarGenes <- head(order(rowVars(assay(vsd)), decreasing = TRUE), 20)
mat  <- assay(vsd)[ topVarGenes, ]
mat  <- mat - rowMeans(mat)
anno <- as.data.frame(colData(vsd)[, c("strain","treatment")])
pheatmap(mat, annotation_col = anno)
```

